Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-07 vesicle_code/docs/adhesion/adhesion.tex
minor changes by yny

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y)$,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to fix the position of the stagnation point by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as particle
image velocimetry of flow inside and around the particle.  To explore
the application of such a fluid trap to measure the adhesion strength
between two bound vesicle membranes, we propose the following thought
experiment.

We begin with two identical vesicles forming a doublet under an
adhesion potential in a quiescent flow as described in
Section~\ref{sec:qflow}.  Once a doublet forms and reaches an
equilibrium configuration, we turn on the fluid trap with the stagnation
point placed at the center of the vesicle doublet.  At low extensional
rates, $\chi$, we expect the vesicle doublet to stay fixed at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

We consider vesicle doublets with reduced areas 0.70, 0.75, 0.80, 0.85,
0.90, and 0.95, all with a length of $\pi$, and we vary the extensional
rate, $\chi$, between $10^{-2}$ and $10^{-1}$.  Since the stagnation
point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the center of the
doublet at each time step so that the stagnation point occurs exactly in
the middle of the doublet.  With this adjustment, the vesicles either
remain as a doublet in the fluid trap centered around the stagnation
point, or the doublet is broken and the vesicles separate from one
another.  In Figures~\ref{fig:extensional1}--\ref{fig:extensional3},
snapshots of a fluid trap containing a vesicle doublet of reduced area
$\Delta A = 0.7$ are simulated at different extensional rates.  While
this doublet remains trapped around the stagnation point for all the
considered extensional rates, a doublet with a reduced area of $\Delta A
= 0.8$ is broken with an extensional rate $\chi = 0.1$
(Figure~\ref{fig:extensional4}).  We summarize the reduced areas and
extensional rates that result in a fluid trap in left plot of
Figure~\ref{fig:extensionalPhaseDiagram}. Parameter values with a blue
mark result in a fluid trap and parameter values with a red mark result
in vesicle separation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y)$,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to place the stagnation point at a desirable location by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as particle
image velocimetry of flow inside and around the particle.  To explore
the application of such a fluid trap to measure the adhesion strength
between two bound vesicle membranes, we propose the following thought
experiment.

We begin with two identical vesicles forming a doublet under an
adhesion potential in a quiescent flow as described in
Section~\ref{sec:qflow}.  Once a doublet forms and reaches an
equilibrium configuration, we turn on the fluid trap with the stagnation
point placed at the center of the vesicle doublet.  At low extensional
rates, we expect the vesicle doublet to stay at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate $\chi_c$ above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate $\chi_c$ on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

@@

We consider vesicle doublets with reduced areas 0.70, 0.75, 0.80, 0.85,
0.90, and 0.95, all with a length of $\pi$, and we vary the extensional
rate, $\chi$, between $10^{-2}$ and $10^{-1}$.  
%
In the simulations we mimic the active control algorithm in the microfluidic experiments by
moving the center of the
doublet at each time step so that the stagnation point occurs exactly in
the middle of the doublet.  
%
With this adjustment, the vesicles either
remain as a doublet in the fluid trap centered around the stagnation
point, or the doublet is broken and the vesicles separate from one
another.  In Figures~\ref{fig:extensional1}--\ref{fig:extensional3},
snapshots of a fluid trap containing a vesicle doublet of reduced area
$\Delta A = 0.7$ are simulated at different extensional rates.  While
this doublet remains trapped around the stagnation point for all the
considered extensional rates, a doublet with a reduced area of $\Delta A
= 0.8$ is broken with an extensional rate $\chi = 0.1$
(Figure~\ref{fig:extensional4}).  We summarize the reduced areas and
extensional rates that result in a fluid trap in left plot of
Figure~\ref{fig:extensionalPhaseDiagram}. Parameter values with a blue
mark result in a fluid trap and parameter values with a red mark result
in vesicle separation.
